---
title: "Simulation - Exponential Distribution"
author: "Segbehoe, Lawrence Sethor"
date: "October 30, 2018"
output:
  pdf_document: default
  html_document:
    df_print: paged
header-includes:
- \usepackage{bbm}
- \usepackage{bm}
- \usepackage{array}
- \usepackage{amssymb}
- \usepackage{amsmath}
- \usepackage{graphicx}
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F, warning = F, fig.width = 11, fig.height = 7, cache = T, eval = T)
## libraries
library(mvtnorm)
library(SpatialExtremes)
library(rmutil)
library(knitr)
```

#### Content

* [Functions](#functions)
  
  * [Function for generating the random numbers](#function-for-generating-the-random-sample)
  
  * [Coverage probability function](#coverage-probability-function)

* [A quick testing of the coverage probability function](#a-quick-testing-of-the-coverage-probability-function)

* [Coverage probability for all pairwise comparisons](#coverage-probability-for-all-pairwise-comparisons)

  * [Normal distribution](#normal-distribution)
  * [Exponential distribution](#exponential-distribution)
  * [Cauchy distribution](#cauchy-distribution)
  * [Laplace distribution](#laplace-distribution)
  * [GEV distribution](#gev-distribution)
  * [Mixture of normal distributions](#mixture-of-normal-distributions)

* [Coverage probability for multiple comparison to control](#coverage-probability-for-multiple-comparison-to-control)

  * [Normal distribution](#normal-distribution)
  * [Exponential distribution](#exponential-distribution)
  * [Cauchy distribution](#cauchy-distribution)
  * [Laplace distribution](#laplace-distribution)
  * [GEV distribution](#gev-distribution)
  * [Mixture of normal distributions](#mixture-of-normal-distribution)







### Functions


#### Function for generating the random sample

* The arguments of the function `sample.fun` are:
   + **n**  = sample size
   + **dist.type** = type of distribution to by used for random sample 

[Back to Content](#content)

```{r}
sample.fun <- function(n, dist.type){
  
  
  # random sample 
    
    if(dist.type == "Normal"|dist.type == "norm"|
               dist.type == "n"|dist.type == "N"|
       dist.type == "Norm"|dist.type == "normal"){
    
          grp1 = rnorm(n, mean = 0, sd = 1)
          grp2 = rnorm(n, mean = 1, sd = 1)
          grp3 = rnorm(n, mean = 2, sd = 1)
        
       dat = cbind(grp1, grp2, grp3)
  
  
    } else if (dist.type == "Exp"|dist.type == "exp"|
               dist.type == "e"|dist.type == "E"){
      
          grp1 = rexp(n, rate = 1)
          grp2 = rexp(n, rate = 2)
          grp3 = rexp(n, rate = 3)
          
      dat = cbind(grp1, grp2, grp3)
      
    } else if (dist.type == "Cauchy"|dist.type == "cauchy"|
               dist.type == "c"|dist.type == "C"){
      
          grp1 = rcauchy(n, location = 0, scale = 1)
          grp2 = rcauchy(n, location = 1, scale = 1)
          grp3 = rcauchy(n, location = 2, scale = 1)
          
      dat = cbind(grp1, grp2, grp3)
          
    } else if (dist.type == "Laplace"|dist.type == "laplace"
               |dist.type == "L"|dist.type == "l"){
      library("rmutil")
      
          grp1 <- rlaplace(n, m = 0, s = 1)
          grp2 <- rlaplace(n, m = 1, s = 1)
          grp3 <- rlaplace(n, m = 2, s = 1)
          
       dat = cbind(grp1, grp2, grp3)
          
    } else if (dist.type == "GEV"|dist.type == "gev"|dist.type == "Gev"|
               dist.type == "g"|dist.type == "G"){
      library("SpatialExtremes")
      
         grp1 <- rgev(n, loc = 0, scale = 1, shape = 0)
         grp2 <- rgev(n, loc = 1, scale = 1, shape = 0)
         grp3 <- rgev(n, loc = 2, scale = 1, shape = 0)
         
     dat = cbind(grp1, grp2, grp3)
    
    } else if(dist.type == "Mixture"|dist.type == "mix"|
               dist.type == "m"|dist.type == "M"|
       dist.type == "normalmixture"|dist.type == "nm"){

      # group 1
        probs = c(.5,.5)          
        dists = runif(n)          
        grp1 = vector(length = n)
      for(i in 1: n){
        if(dists[i] < probs[1]){
          grp1[i] = rnorm(1, mean=0, sd=1)
        } else {
          grp1[i] = rnorm(1, mean=1, sd=1)
        }
      }
        
        # group 2
        probs = c(.5,.5)          
        dists = runif(n)          
        grp2 = vector(length = n)
      for(i in 1: n){
        if(dists[i] < probs[1]){
          grp2[i] = rnorm(1, mean=1, sd=1)
        } else {
          grp2[i] = rnorm(1, mean=2, sd=1)
        }
      }
        
        # group 3
        probs = c(.5,.5)          
        dists = runif(n)          
        grp3 = vector(length = n)
      for(i in 1: n){
        if(dists[i] < probs[1]){
          grp3[i] = rnorm(1, mean=2, sd=1)
        } else {
          grp3[i] = rnorm(1, mean=3, sd=1)
        }
      }
        
      dat = cbind(grp1, grp2, grp3)
      
  }
  
  
  
  return(dat)
}


## testing the sample.fun functions
sample.fun(10, "g")
```



#### Coverage probability function


* The arguments of the function `Coverage.Prob` are
    + **R**  = number of replications
    + **n**  = sample size
    + **comp.type**  = type of comparison to use with options
      + **ap** = all pairwise comparison and 
      + **mc** = multiple comparison to control
    + **dist.type**  = type of distribution to be used for the random samples

[Back to Content](#content)


```{r}

Coverage.Prob <- function(R = 1000, n, comp.type , dist.type ){


      R  #replications
      n  # sample size
      
      k = 3 # number of  groups
      grp = 1:k # groups
      
      # container for density estimates at the median
      est.f.md = vector(length = k )
      
      ci = vector(length = R)
      
      # dat = NULL


    for (i in 1:R) {
      
      # get sample for the three groups
      dat = sample.fun(n, dist.type)
      
      # get the median of each group
      md = apply(dat, 2, median)
      
      
      if(comp.type == "ap"){
  
      # design matrix for all pairwise comparison
      design.matrix <- matrix(c(1,-1, 0,1,0,-1,0,1,-1), nrow = 3, byrow = T)
      
      } else if (comp.type == "mc"){
        
      # design matrix for multiple comparison to control
      design.matrix <- matrix(c(1,0,-1,0,1,-1), nrow = 2, byrow = T)
      
      }
      
      
      # get delta
      DELTA = design.matrix%*%md   # all pairwise comparison
      
      
      # True DELTA 
      # It depends on the distribution being used to for the random sample
      
    if(dist.type == "Normal"|dist.type == "norm"|
               dist.type == "n"|dist.type == "N"|
       dist.type == "Norm"|dist.type == "normal"){
    
      true.DELTA = design.matrix%*%c(0,1,2)
      
    } else if (dist.type == "Exp"|dist.type == "exp"|
               dist.type == "e"|dist.type == "E"){
      
      true.DELTA = design.matrix%*%c(log(2), 0.5*log(2), (1/3)*log(2))
     
    } else if (dist.type == "Cauchy"|dist.type == "cauchy"|
               dist.type == "c"|dist.type == "C"){
      
      true.DELTA = design.matrix%*%c(0,1,2)
    
    } else if (dist.type == "Laplace"|dist.type == "laplace"
               |dist.type == "L"|dist.type == "l"){
      
      true.DELTA = design.matrix%*%c(0,1,2)
    
    } else if (dist.type == "GEV"|dist.type == "gev"|dist.type == "Gev"|
               dist.type == "g"|dist.type == "G"){
      
      true.DELTA = design.matrix%*%c(-log(log(2)), 1- log(log(2)), 2 -log(log(2)))
    
    } else if (dist.type == "Mixture"|dist.type == "mix"|
               dist.type == "m"|dist.type == "M"|
       dist.type == "normalmixture"|dist.type == "nm"){
   
      true.DELTA = design.matrix%*%c((0.5),(3/2),(5/2))
    
    } 
      
     
      ## density estimates at the median
      
      
          # treatment 1
          den.grp1 <- density(dat[,1])
          datfram.grp1 <- data.frame(x = den.grp1$x, y = den.grp1$y)
          ix1 = which.min(abs(den.grp1$x - md[grp[1]]))
          est.f.md[grp[1]] <- datfram.grp1[ix1,"y"]
          
          # treatment 2
          den.grp2 <- density(dat[,2])
          datfram.grp2 <- data.frame(x = den.grp2$x, y = den.grp2$y)
          ix2 = which.min(abs(den.grp2$x - md[grp[2]]))
          est.f.md[grp[2]] <- datfram.grp2[ix2,"y"]
          
          # control 1
           den.grp3 <- density(dat[,3])
          datfram.grp3 <- data.frame(x = den.grp3$x, y = den.grp3$y)
          ix3 = which.min(abs(den.grp3$x - md[grp[3]]))
          est.f.md[grp[3]] <- datfram.grp3[ix3,"y"]
    
    
    # Compute the Confidence intervals      
          
      # vector of variance 1 for each group
      sig2 <- 1/(4*n*(est.f.md^2))
      
      # get the SIGMA
      SIGMA <- diag(sig2)
    
      # covariance matrix
      cov.mat <- design.matrix%*%SIGMA%*%t(design.matrix)
    
      # building the correlation matrix associated with the
      # the covariance matrix
      D <- sqrt(diag(diag(cov.mat)))
      D.inv <- solve(D)
      corr.mat <- D.inv%*%cov.mat%*%D.inv
    
      # equicoordinate quantile 
      qq <- qmvnorm(0.95, sigma = corr.mat, tail = "both")$quantile
    
      C <- qq*sqrt(diag(cov.mat))
      
      # confidence interval
      CI = cbind(DELTA - C, DELTA + C)
      
      
      if(comp.type == "ap"){
  
       ci[i] =  ((CI[1,1] < true.DELTA[1,1]) & (true.DELTA[1,1] < CI[1,2] ) &
               (CI[2,1] < true.DELTA[2,1]) & (true.DELTA[2,1] < CI[2,2] ) &
               (CI[3,1] < true.DELTA[3,1]) & (true.DELTA[3,1] < CI[3,2] ) )
           
      } else if (comp.type == "mc"){
        
        ci[i] = ((CI[1,1] < true.DELTA[1,1]) & (CI[1,2] > true.DELTA[1,1]) &
                (CI[2,1] < true.DELTA[2,1]) & (CI[2,2] > true.DELTA[2,1]))
      }
     
      
      
    }

  CP = mean(ci)
  return(CP)

}


```


## A quick testing of the coverage probability function

  * For 100 replications 
  * sample size = 100 
  * all pairwise comparison
  * For mixture of normal distribution

[Back to Content](#content)
```{r ,eval=FALSE}
# for 100 replications and sample size = 100 and
# all pairwise comparison for mixture distribution
CP = Coverage.Prob(R = 100, n =100, comp.type = "ap", dist.type = "Mixture")
data.frame(Coverage_Prob = CP)
```



\newpage

## Coverage probability for all pairwise comparisons

  * For a quick calculations and check on the code, I do the simulation for just **100** replications for all distributions under all pairwise comparisons
  
  
[Back to Content](#content)


### Normal distribution
```{r,eval=FALSE}

N = c(10, 30, 100, 500)

CP = vector(length = length(N))

for (i in 1:length(N)) {
  sample.size = N[i]
  CP[i] = Coverage.Prob(R = 100, n = 20, comp.type = "ap", dist.type = "normal")
}

Cov.Prob = data.frame(n = N, Coverage.Probability = CP)


kable(Cov.Prob, caption = c(paste("Coverage Probabilities by sample"),
                         paste("All pairwise Comparisons - 
                               Normal Distribution")))

```

### Exponential distribution


[Back to Content](#content)
```{r,eval=FALSE}

N = c(10, 30, 100, 500)

CP = vector(length = length(N))

for (i in 1:length(N)) {
  n = N[i]
  CP[i] = Coverage.Prob(R = 100, n = n, comp.type = "ap", dist.type = "exp")
}

Cov.Prob = data.frame(n = N, Coverage.Probability = CP)


kable(Cov.Prob, caption = c(paste("Coverage Probabilities by sample"),
                         paste("All pairwise Comparisons - 
                               Exponential Distribution")))

```


### Cauchy distribution

[Back to Content](#content)

```{r,eval=FALSE}

N = c(10, 30, 100, 500)

CP = vector(length = length(N))

for (i in 1:length(N)) {
  n = N[i]
  CP[i] = Coverage.Prob(R = 100, n = n, comp.type = "ap", dist.type = "Cauchy")
}

Cov.Prob = data.frame(n = N, Coverage.Probability = CP)


kable(Cov.Prob, caption = c(paste("Coverage Probabilities by sample"),
                         paste("All pairwise Comparisons - 
                               Cauchy Distribution")))

```




### Laplace distribution


[Back to Content](#content)

```{r,eval=FALSE}

N = c(10, 30, 100, 500)

CP = vector(length = length(N))

for (i in 1:length(N)) {
  n = N[i]
  CP[i] = Coverage.Prob(R = 100, n = n, comp.type = "ap", dist.type = "Laplace")
}

Cov.Prob = data.frame(n = N, Coverage.Probability = CP)


kable(Cov.Prob, caption = c(paste("Coverage Probabilities by sample"),
                         paste("All pairwise Comparisons - 
                               Laplace Distribution")))

```



### GEV distribution

[Back to Content](#content)

```{r,eval=FALSE}

N = c(10, 30, 100, 500)

CP = vector(length = length(N))

for (i in 1:length(N)) {
  n = N[i]
  CP[i] = Coverage.Prob(R = 100, n = n, comp.type = "ap", dist.type = "GEV")
}

Cov.Prob = data.frame(n = N, Coverage.Probability = CP)


kable(Cov.Prob, caption = c(paste("Coverage Probabilities by sample"),
                         paste("All pairwise Comparisons - 
                               GEV Distribution")))

```




### Mixture of normal distributions


[Back to Content](#content)

```{r,eval=FALSE}

N = c(10, 30, 100, 500)

CP = vector(length = length(N))

for (i in 1:length(N)) {
  n = N[i]
  CP[i] = Coverage.Prob(R = 100, n = n, comp.type = "ap", dist.type = "Mixture")
}

Cov.Prob = data.frame(n = N, Coverage.Probability = CP)


kable(Cov.Prob, caption = c(paste("Coverage Probabilities by sample"),
                         paste("All pairwise Comparisons - 
                               Mixture of normal distribution")))

```


\newpage
## Coverage probability for multiple comparison to control

* For a quick calculations and check on the code, I do the simulation for just **100** replications for all distributions under multiple comparison to control
  


### Normal distribution


[Back to Content](#content)

```{r,eval=FALSE}

N = c(10, 30, 100, 500)

CP = vector(length = length(N))

for (i in 1:length(N)) {
  n = N[i]
  CP[i] = Coverage.Prob(R = 100, n = n, comp.type = "mc", dist.type = "normal")
}

Cov.Prob = data.frame(n = N, Coverage.Probability = CP)


kable(Cov.Prob, caption = c(paste("Coverage Probabilities by sample"),
                         paste("Multiple comparison to control - 
                               Normal Distribution")))

```

### Exponential distribution


[Back to Content](#content)

```{r,eval=FALSE}

N = c(10, 30, 100, 500)

CP = vector(length = length(N))

for (i in 1:length(N)) {
  n = N[i]
  CP[i] = Coverage.Prob(R = 100, n = n, comp.type = "mc", dist.type = "exp")
}

Cov.Prob = data.frame(n = N, Coverage.Probability = CP)


kable(Cov.Prob, caption = c(paste("Coverage Probabilities by sample"),
                         paste("Multiple comparison to control - 
                               Exponential Distribution")))

```


### Cauchy distribution


[Back to Content](#content)

```{r,eval=FALSE}

N = c(10, 30, 100, 500)

CP = vector(length = length(N))

for (i in 1:length(N)) {
  n = N[i]
  CP[i] = Coverage.Prob(R = 100, n = n, comp.type = "mc", dist.type = "Cauchy")
}

Cov.Prob = data.frame(n = N, Coverage.Probability = CP)


kable(Cov.Prob, caption = c(paste("Coverage Probabilities by sample"),
                         paste("Multiple comparison to control - 
                               Cauchy Distribution")))

```




### Laplace distribution


[Back to Content](#content)

```{r,eval=FALSE}

N = c(10, 30, 100, 500)

CP = vector(length = length(N))

for (i in 1:length(N)) {
  n = N[i]
  CP[i] = Coverage.Prob(R = 100, n = n, comp.type = "mc", dist.type = "Laplace")
}

Cov.Prob = data.frame(n = N, Coverage.Probability = CP)


kable(Cov.Prob, caption = c(paste("Coverage Probabilities by sample"),
                         paste("Multiple comparison to control - 
                               Laplace Distribution")))

```



### GEV distribution


[Back to Content](#content)

```{r,eval=FALSE}

N = c(10, 30, 100, 500)

CP = vector(length = length(N))

for (i in 1:length(N)) {
  n = N[i]
  CP[i] = Coverage.Prob(R = 100, n = n, comp.type = "mc", dist.type = "GEV")
}

Cov.Prob = data.frame(n = N, Coverage.Probability = CP)


kable(Cov.Prob, caption = c(paste("Coverage Probabilities by sample"),
                         paste("Multiple comparison to control - 
                               GEV Distribution")))

```




### Mixture of normal distribution


[Back to Content](#content)


```{r ,eval=FALSE}

N = c(10, 30, 100, 500)

CP = vector(length = length(N))

for (i in 1:length(N)) {
  n = N[i]
  CP[i] = Coverage.Prob(R = 100, n = n, comp.type = "mc", dist.type = "Mixture")
}

Cov.Prob = data.frame(n = N, Coverage.Probability = CP)


kable(Cov.Prob, caption = c(paste("Coverage Probabilities by sample"),
                         paste("Multiple comparison to control - 
                               Mixture of normal distributions")))

```


[Back to Content](#content)






## Histogram of the true density and estimated density

[Back to Content](#content)

  * For Cauchy(0,1) the true density and the estimated densities are shown by sample size.
  * From the 10000 replications, the second row shows the histogram of $\hat{f}(\hat{\xi})$ and the dotted vertical lines represent $f(\xi)$

```{r , fig.height=6, fig.width=11, eval= T}

f <- function(n){seq(-10, 10, length.out = n)}

histogram.fun <- function(R, dist.type){


  # container for density estimates at the median
      est.f.md5 = NULL
      est.f.md10 = NULL
      est.f.md100 = NULL
      est.f.md500 = NULL
  
      # groups
      grp = 1:3
 
      k = 1
       for(i in 1 :R){
    
    dat_n5 <- sample.fun(5, dist.type)
    dat_n10 <- sample.fun(10, dist.type)
    dat_n100 <- sample.fun(100, dist.type)
    dat_n500 <- sample.fun(500, dist.type)
    
    
    md_n5 = apply(dat_n5, 2, median)
    md_n10 = apply(dat_n10, 2, median)
    md_n100 = apply(dat_n100, 2, median)
    md_n500 = apply(dat_n500, 2, median)
    
    ## density estimates at the median
      
          # n =5
          den.grp1 <- density(dat_n5[,1])
          datfram.grp1 <- data.frame(x = den.grp1$x, y = den.grp1$y)
          ix1 = which.min(abs(den.grp1$x - md_n5[grp[1]]))
          est.f.md5[k] <- datfram.grp1[ix1,"y"]
          
          # n =10
          den.grp1 <- density(dat_n10[,1])
          datfram.grp1 <- data.frame(x = den.grp1$x, y = den.grp1$y)
          ix1 = which.min(abs(den.grp1$x - md_n10[grp[1]]))
          est.f.md10[k] <- datfram.grp1[ix1,"y"]
          
          # n = 100
           den.grp1 <- density(dat_n100[,1])
          datfram.grp1 <- data.frame(x = den.grp1$x, y = den.grp1$y)
          ix1 = which.min(abs(den.grp1$x - md_n100[grp[1]]))
          est.f.md100[k] <- datfram.grp1[ix1,"y"]
          
          # n = 500
           den.grp1 <- density(dat_n500[,1])
          datfram.grp1 <- data.frame(x = den.grp1$x, y = den.grp1$y)
          ix1 = which.min(abs(den.grp1$x - md_n500[grp[1]]))
          est.f.md500[k] <- datfram.grp1[ix1,"y"]
          
      k = k +1   
      
    }
 
  dat_n5 <- sample.fun(5, dist.type)
  dat_n10 <- sample.fun(10, dist.type)
  dat_n100 <- sample.fun(100, dist.type)
  dat_n500 <- sample.fun(500, dist.type)      
           


if (dist.type == "Cauchy"|dist.type == "cauchy"|
               dist.type == "c"|dist.type == "C"){
par(mfrow = c(2, 4))  
 plot(density(dat_n5[, 1]), main = "",  xlim = c(-10, 10), ylim = c(0,.4))
lines(f(512), dcauchy(f(512)), lty = 3, col = 4, lwd = 2)
plot(density(dat_n10[, 1]), main = "", ylim = c(0,.4),  xlim = c(-10, 10))
lines(f(512), dcauchy(f(512)), lty = 3, col = 4, lwd = 2)
plot(density(dat_n100[, 1]), main = "", xlim = c(-10, 10), ylim = c(0,.4) )
lines(f(512), dcauchy(f(512)), lty = 3, col = 4, lwd = 2)
plot(density(dat_n500[, 1]), main = "", xlim = c(-20, 20), ylim = c(0,.4))
lines(f(512), dcauchy(f(512)), lty = 3, col = 4, lwd = 2)


hist(est.f.md5, main = "N = 5", freq = F,breaks = "fd",xlim = c(0,2.0),
     las =1, xlab = expression(hat(italic(f))(hat(xi))))
abline(v = (1/pi), col = 2, lty = 2, lwd = 2)

hist(est.f.md10, nclass = 60, main = "N = 10", freq = F,
     las =1, xlab = expression(hat(f)(hat(xi))))
abline(v = (1/pi), col = 2, lty = 2, lwd = 2)


hist(est.f.md100, nclass = 60, main = "N = 100", freq = F,
     las =1, xlab = expression(hat(f)(hat(xi))))
abline(v = (1/pi), col = 2, lty = 2, lwd = 2)

hist(est.f.md500, nclass = 60, main = "N = 500", freq = F,
     las =1, xlab = expression(hat(f)(hat(xi))))
abline(v =( 1/pi), col = 2, lty = 2, lwd = 2)    
    


par(mfrow = c(1,1))

    } else if (dist.type == "Laplace"|dist.type == "laplace"
               |dist.type == "L"|dist.type == "l"){
      
      library(rmutil)
par(mfrow = c(2, 4))

 plot(density(dat_n5[, 1]), main = "",  xlim = c(-10, 10), ylim = c(0,.5))
lines(f(512), dlaplace(f(512)), lty = 3, col = 4, lwd = 2)
plot(density(dat_n10[, 1]), main = "", ylim = c(0,.5),  xlim = c(-10, 10))
lines(f(512), dlaplace(f(512)), lty = 3, col = 4, lwd = 2)
plot(density(dat_n100[, 1]), main = "" , ylim = c(0,.5), xlim = c(-10, 10))
lines(f(512), dlaplace(f(512)), lty = 3, col = 4, lwd = 2)
plot(density(dat_n500[, 1]), main = "", xlim = c(-10, 10), ylim = c(0,.5))
lines(f(512), dlaplace(f(512)), lty = 3, col = 4, lwd = 2) 



hist(est.f.md5, main = "N = 5", freq = F,breaks = "fd",xlim = c(0,2.0),
     las =1, xlab = expression(hat(italic(f))(hat(xi))))
abline(v = (1/2), col = 2, lty = 2, lwd = 2)

hist(est.f.md10, nclass = 60, main = "N = 10", freq = F,
     las =1, xlab = expression(hat(f)(hat(xi))))
abline(v = (1/2), col = 2, lty = 2, lwd = 2)

hist(est.f.md100, nclass = 60, main = "N = 100", freq = F,
     las =1, xlab = expression(hat(f)(hat(xi))))
abline(v = (1/2), col = 2, lty = 2, lwd = 2)

hist(est.f.md500, nclass = 60, main = "N = 500", freq = F,
     las =1, xlab = expression(hat(f)(hat(xi))))
abline(v =( 1/2), col = 2, lty = 2, lwd = 2)


par(mfrow = c(1,1))

}


}
```







```{r}

histogram.fun(R = 10000, "c")



```





```{r}

histogram.fun(R = 10000, "l")



```



